The connection between star formation rate and dark matter halo mass in the epoch of reionization

### 2015-04-29 paper/ms.tex
Abstract. (+Word Wrap)

\keywords{galaxies: high-redshift --- methods: numerical} 
\begin{abstract}
We present updated constraints on the relationship between the star
formation rate and dark matter halo mass at redshift $z\sim 6$.
The observational basis for our work is the restframe UV luminosity
function data obtained with CFHTLS, HST Legacy Survey and UKIDSS.
The constraints are based on an abundance matching methodology to the
observational data using cosmological N-body simulations.
We also take into account the influence on the results of the dust extinction
scaling derived from observations by Bouwens.
We compare our results against the results of abundance matching
methods (to the stellar mass), a semi-analytic model of galaxy
formation (GALFORM) and a hydrodynamical simulation (Illustris).
\end{abstract}

@@

\citet{McLure09} build the luminosity function for $z=5$ and $z=6$ using data from two ground-based
 telescopes: the United Kingdom Infrared Telescope in the near-IR imaging and Subaru 
 Telescope for the optical imaging. They use the first data release of the UKIRT Infrared 
 DeepSky Survey Ultra Deep Survey (UDS), together with the Subaru XMM-Newton 
 Survey (SXDS). Total observed area is $0.63 \deg^2$ uniformly covered by booth catalogues.
The volume in this survey is $\sim 3 \times 10^6 \textrm{Mpc}^3$.
The UKIRT was equipped with the WFCAM using $J K$ filters. The Subaru was equipped 
with the Suprime-Cam with the $B V R i' z'$ filters. All candidates where brighter than 
$z'=26$. The UV rest frame magnitude range is $-22.4\leq M_{1500} \leq-20.6$. The LF 
was calculated using the maximum likelihood estimator of \citet{Schmidt68}. 
Their analysis gave a total number of 104 LBG candidates in the redshift range 
$5.7\leq z \leq 6.3$. LF was parameterized according to the Schechter function with 
$\phi^*  =(1.8 \pm 0.5) \times 10 ^{-3}  \textrm{Mpc}^{-3}$, $M^*_{1500} = -20.04\pm 0.12$ 
and $\alpha = -1.71 \pm 0.11$.

@@

\subsection{MCMC}
We used a Markov Chain Monte Carlo method to find the best parameters and its 
uncertainties over each one of the boxes and each observational dataset. The code was written
on Python using the Scipy, Numpy and Matplotlib libraries.

@@

We performed 10.000 effective MCMC steps, plus 1.000 thermalization steps over each box.
We repeated for the same box without consider dust attenuation.

@@

Finally, the UV luminosity for each galaxy can be directly related with SFR according 
to \citet{Madau98}.
This model is accurate within the range of $10^8 - 10^9 \textrm{yr}$\citep{Kennicutt98}.
The relation between UV luminosity and Star Formation Rate \citep{Madau98} and \cite{Kennicutt98} 
is given by:
\begin{equation}
 \textrm{SFR}\left(M_\odot \textrm{yr}^{-1}\right) 
      = 1.4 \times 10^{-28} L_{\nu} \left( \textrm{erg s}^{-1}\textrm{Hz}^{-1} 
	\right)
\end{equation}

@@

In this model we have made two assumptions:
\begin{enumerate}
 \item Each halo in the catalog hosts one galaxy. There are not empty
halos, also none of halos has two or more galaxies.
 \item The UV luminosity of each galaxy is function of one variable: the mass of
the DMH in wich is located.
\end{enumerate}

The simplest relation we can have is a powerlaw:
 \begin{equation}
  L = L_0 M^\alpha
 \end{equation}
but has not well agreement with observed luminosity functions.

A better model is a four parameter function. Each galaxy has a luminosity given
by:
  \begin{equation}
  L = L_{0} M \left[ \left( \frac{M}{M_0}\right)^{-\beta} 
		   + \left( \frac{M}{M_0}\right)^{\gamma} 
               \right]^{-1}
  \end{equation}
where $M$ is the hosting DMH mass, $L_{0}$ is a normalization constant, $M_0$
is the critical mass where the luminosity function has a slope change, 
$\beta$ and $\gamma$ are the slopes. This equation has a similar fashion to the
mass to light relation \citep{vandenBosch03} and the mean relation between
stellar mas of a galaxy and the mass of its halo used by \cite{Moster10}.

There are more complex models\citep{Lee09} that includes a random behavior:
galaxies has not synchronization on the beginning of star forming stage, also
this stage may be time limited. This is called duty cycle. It is probable to
have in the observations some invisible galaxies in the UV continuum due their
duty cycle may has not started as well it may ended. Also may be present a
normal distribution of the luminosity around the expected values.

@@

\begin{table}
\begin{center}
\begin{tabular}{ccccc}\hline\hline
Parameter  & Mean value & Cosmic Variance & MCMC error\\\hline
$\log_{10}(L_0)$   & 18.067   & 0.029   & 0.116 \\
$\log_{10}(M_0)$  & 11.190   & 0.072   & 0.729 \\
$\beta$                  & 1.4801  & 0.175   & 1.299 \\
$\gamma$             & 0.40284 & 0.058  & 0.3147\\
\hline\hline
\end{tabular}
\caption{  Average Best fit parameters to the Willott data in the Dust Attenuation Model.} 
\label{table:Willott_best_fit_parameters}
\end{center}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{ccccc}\hline\hline
Parameter  & Mean value & Cosmic Variance & MCMC error\\\hline
$\log_{10}(L_0)$   & 17.807   & 0.020   & 0.150 \\
$\log_{10}(M_0)$  & 11.119    & 0.060   & 0.642 \\
$\beta$                  &  1.437    & 0.178   & 1.374 \\
$\gamma$             &   0.518   & 0.046  & 0.246\\
\hline\hline
\end{tabular}
\caption{  Average Best fit parameters to the Willott data in the No-Dust Attenuation Model.} 
\label{table:Willott_best_fit_parameters_no_dust}
\end{center}
\end{table}

@@

\begin{figure*}
\epsscale{1.00}
\plotone{fig/OD1_w_dust_box_0.png} % pdf available
\caption{Parameter dispersion fitting the Willott with the Dust Attenuation model. 
$1\sigma$ is defined by the likelihood ratio between 0.0 (red) and 0.5 (green)}
\label{fig:OD1_MCMC_best_steps_w}
\end{figure*}

\begin{figure*}
\epsscale{1.00}
\plotone{fig/OD1_wo_dust_box_0.png} % pdf available
\caption{Parameter dispersion fitting the Willott with the No-Dust Attenuation model. 
$1\sigma$ is defined by the likelihood ratio between 0.0 (red) and 0.5 (green)}
\label{fig:OD1_MCMC_best_steps_wo}
\end{figure*}

@@

\begin{figure}
\epsscale{1.00}
\plotone{fig/SFR_DMHM_wo.pdf}
\caption{Star formation rate as function of the dark matter halo mass without dust attenuation. 
Solid lines represents the mean SFR value over the small boxes within $50\%$  shaded region. 
Comparison with the GALFORM semi-analitical model \citep{Gonzalez14} and a implementation
of abundance matching model \citep{Behroozi13}. }
\label{graph_SFR_DMHM_wo}
\end{figure}

\begin{figure}
\epsscale{1.00}
\plotone{fig/SFR_DMHM_w.pdf}
\caption{Star formation rate as function of the dark matter halo mass with dust attenuation. 
Solid lines represents the mean SFR value over the small boxes within $50\%$  shaded region.
Comparison with the GALFORM semi-analitical model \citep{Gonzalez14} and a implementation
of abundance matching model \citep{Behroozi13}. }
\label{graph_SFR_DMHM_w}
\end{figure}

@@

>>>>

\keywords{galaxies: high-redshift --- methods: numerical} 
\begin{abstract}
We present updated constraints on the relationship between the star
formation rate and dark matter halo mass at redshift $z\sim 6$.
The observational basis for our work is the restframe UV luminosity
function data obtained with CFHTLS, HST Legacy Survey and UKIDSS.
The constraints are based on an abundance matching methodology to the
observational data using cosmological N-body simulations.
We also take into account the influence on the results of the dust extinction
scaling derived from observations by Bouwens.
We show dependence of our parameter uncertainties related to the
comoving volume and magnitude range in observed surveys, taking advantage of the 
wide mass resolution range of latest available dark matter halo catalogs.
We compare our results against the results of abundance matching
methods (to the stellar mass), a semi-analytic model of galaxy
formation (GALFORM) and a hydrodynamical simulation (Illustris).
\end{abstract}

@@

\citet{McLure09} build the luminosity function for $z=5$ and $z=6$ using data 
from two ground-based telescopes: the United Kingdom Infrared Telescope in the 
near-IR imaging and Subaru Telescope for the optical imaging. They use the first 
data release of the UKIRT Infrared  DeepSky Survey Ultra Deep Survey (UDS), 
together with the Subaru XMM-Newton  Survey (SXDS). Total observed area is 
$0.63 \deg^2$ uniformly covered by booth catalogues.
The volume in this survey is $\sim 3 \times 10^6 \textrm{Mpc}^3$.
The UKIRT was equipped with the WFCAM using $J K$ filters. The Subaru was 
equipped with the Suprime-Cam with the $B V R i' z'$ filters. All candidates 
where brighter than $z'=26$. The UV rest frame magnitude range is 
$-22.4\leq M_{1500} \leq-20.6$. The LF was calculated using the maximum 
likelihood estimator of \citet{Schmidt68}. Their analysis gave a total number of 
104 LBG candidates in the redshift range $5.7\leq z \leq 6.3$. LF was 
parameterized according to the Schechter function with 
$\phi^* =(1.8\pm 0.5)\times 10 ^{-3} \textrm{Mpc}^{-3}$, 
$M^*_{1500} = -20.04\pm 0.12$ and $\alpha = -1.71 \pm 0.11$.

@@

\subsection{MCMC}
We used a Markov Chain Monte Carlo method to find the best parameters and its 
uncertainties over each one of the boxes and each observational dataset. The 
code was written on Python using the SciPy Library \citep{SciPy}
the IPython environment \citep{IPython}.

@@

We performed 100.000 effective MCMC steps, plus 10.000 thermalization steps over 
each box. We repeated for the same box without consider dust attenuation.

@@

Finally, the UV luminosity for each galaxy can be directly related with SFR according 
to \citet{Madau98}.
This model is accurate within the range of $10^8 - 10^9 \textrm{yr}$
\citep{Kennicutt98}. The relation between UV luminosity and Star Formation Rate 
is given by:
\begin{equation}
 \textrm{SFR}\left(M_\odot \textrm{yr}^{-1}\right) 
      = 1.4 \times 10^{-28} L_{\nu} \left( \textrm{erg s}^{-1}\textrm{Hz}^{-1} 
	\right)
\end{equation}

@@

In this model we have made two assumptions:
\begin{enumerate}
 \item Each halo in the catalog hosts one galaxy. There are not empty
halos, also none of halos has two or more galaxies.
 \item The UV luminosity of each galaxy is function of one variable: the mass of
the DMH in wich is located.
\end{enumerate}
Our model is a four parameter function. Each galaxy has a luminosity given
by:
  \begin{equation}
  L = L_{0} M \left[ \left( \frac{M}{M_0}\right)^{-\beta} 
		   + \left( \frac{M}{M_0}\right)^{\gamma} 
               \right]^{-1}
  \end{equation}
where $M$ is the hosting DMH mass, $L_{0}$ is a normalization constant, $M_0$
is the critical mass where the luminosity function has a slope change, 
$\beta$ and $\gamma$ are the slopes. This equation has a similar fashion to the
mass to light relation \citep{vandenBosch03} and the mean relation between
stellar mas of a galaxy and the mass of its halo used by \cite{Moster10}.

There exists more sophisticated models as \cite{Lee09} shown, with scattering in 
the DM Mass-Stellar Mass, but also including some 
randomization in star forming histories, starburst episodes (duty cycle); 
galaxies have not synchronization on the beginning of star forming stage, also
this stage may be time limited. It results in the existence of massive galaxies
in the studied volume, but with no emmission in the UV continuum due their
duty cycle may has not started as well it may ended. Also may be present a
normal distribution of the luminosity around the expected values.

@@

\begin{table}
\begin{center}
\begin{tabular}{cccccc}\hline\hline
 & \multicolumn{2}{c}{Dust Att.} & \multicolumn{2}{c}{No-Dust Att.}\\
Parameter                  & MCMC                  & C.Var. & MCMC                    & C.Var. \\\hline
$\log_{10}(L_0/L_{\odot})$ & $18.07_{-0.06}^{+0.10}$ & 0.03 & $17.81_{-0.08}^{+0.12}$ & 0.02  \\
$\log_{10}(M_0/M_{\odot})$ & $11.19_{-0.03}^{+0.67}$ & 0.07 & $11.12_{-0.24}^{+0.58}$ & 0.06  \\
$\beta$                    & $1.480_{-1.28}^{+0.11}$ & 0.18 & $ 1.44_{-1.35}^{+0.16}$ & 0.18  \\
$\gamma\times0.1$          & $4.028_{-0.82}^{+3.01}$ & 0.58 & $ 5.18_{-0.59}^{+2.36}$ & 0.46  \\
\hline\hline
\end{tabular}
\caption{ Best fit parameters to the Willott data with and without 
dust attenuation over 64 small boxes. Mean best value estimated 
with MCMC, mean $1\sigma$ confidence interval and Cosmic Variance
(C.Var.)
} 
\label{table:Willott_best_fit_parameters_no_dust}
\end{center}
\end{table}

@@

\begin{figure}
\epsscale{1.00}
\plotone{fig/OD1_w_dust_box_0.png} % pdf available
\caption{Parameter dispersion fitting the Willott with the Dust Attenuation model. 
$1\sigma$ is defined by the likelihood ratio between 0.0 (red) and 0.5 (green)}
\label{fig:OD1_MCMC_best_steps_w}
\end{figure}

\begin{figure}
\epsscale{1.00}
\plotone{fig/OD1_wo_dust_box_0.png} % pdf available
\caption{Parameter dispersion fitting the Willott with the No-Dust Attenuation model. 
$1\sigma$ is defined by the likelihood ratio between 0.0 (red) and 0.5 (green)}
\label{fig:OD1_MCMC_best_steps_wo}
\end{figure}

@@

\begin{figure}
\epsscale{1.00}
\plotone{fig/SFR_DMHM_wo.pdf}
\caption{Star formation rate as function of the dark matter halo mass without dust attenuation. 
Solid lines represents the mean SFR value over the small boxes within $50\%$  shaded region. 
Comparison with the GALFORM semi-analitical model \citep{Gonzalez14} and a implementation
of abundance matching model \citep{Behroozi13}. }
\label{fig:SFR_DMHM_wo}
\end{figure}

\begin{figure}
\epsscale{1.00}
\plotone{fig/SFR_DMHM_w.pdf}
\caption{Star formation rate as function of the dark matter halo mass with dust attenuation. 
Solid lines represents the mean SFR value over the small boxes within $50\%$  shaded region.
Comparison with the GALFORM semi-analitical model \citep{Gonzalez14} and a implementation
of abundance matching model \citep{Behroozi13}. }
\label{fig:_SFR_DMHM_w}
\end{figure}

@@

\begin{table*}
\begin{center}
\begin{tabular}{cccccc}\hline\hline
 & \multicolumn{2}{c}{Dust Attenuation} & \multicolumn{2}{c}{No-Dust Attenuation}\\
Parameter  & MCMC & Cosmic Variance & MCMC & Cosmic Variance \\\hline
$\log_{10}(L_0)$   & $18.067_{-0.059}^{+0.096}$   & 0.029 & $17.807_{-0.083}^{+0.115}$   & 0.020  \\
$\log_{10}(M_0)$   & $11.190_{-0.026}^{+0.674}$   & 0.072 & $11.119_{-0.237}^{+0.582}$   & 0.060  \\
$\beta$            & $1.4801_{-1.276}^{+0.114}$   & 0.175 & $ 1.437_{-1.348}^{+0.157}$   & 0.178  \\
$\gamma\times0.1$  & $4.0284_{-0.822}^{+3.010}$   & 0.576 & $ 5.182_{-0.588}^{+2.366}$  & 0.461  \\
\hline\hline
\end{tabular}
\caption{  Average Best fit parameters to the Willott data with and without dust attenuation.} 
\label{table:Willott_best_fit_parameters_no_dust}
\end{center}
\end{table*}
